### 2018-04-05 paper/user-model.tex

We will first present a model of each user's behavior, that will drive the connection intervals of our system's nodes.

@@

In our case, the states are the different locations where the user is susceptible to go.
We considered the three following locations: $\mathcal{S}=\left\{ \mathit{work}, \mathit{outside}, \mathit{home} \right\}$. 
There is one action per device $d\in\mathcal{D}$, plus a ``none'' action when the user does not do anything: 
$\mathcal{O} = \left\{ t_d \right\}_{d\in \mathcal{D}} \bigcup \left\{ t_\text{none} \right\}$.
$t_d$ means that the user turned her device $d$ on or off (i.e. \emph{toggled} it), depending on the device's previous state.

@@

$$ \forall d \in \mathcal{D},
\begin{cases}
c^{(d)}_t=0 &\text{when } t=0; \\
c^{(d)}_t=c^{(d)}_{t-1} \oplus \left[ O_t = t_d \right] &\text{when } t>0.
\end{cases}$$

@@

>>>>

We will first present a model of each user's behavior, that will drive the connection intervals of our system's nodes.
Then, we will show how devices can predict the future user's behavior given the model.

@@

In our case, the states are the different locations where the user is susceptible to go.
We considered the three following locations: $\mathcal{S}=\left\{ \mathit{work}, \mathit{outside}, \mathit{home} \right\}$. 
There is one action per device $d\in\mathcal{D}$, plus a ``none'' action when the user does not do anything: 
$\mathcal{O} = \left\{ t_d \right\}_{d\in \mathcal{D}} \bigcup \left\{ t_\text{none} \right\}$.
$t_d$ means that the user turned her device $d$ on or off (i.e. \emph{toggled} it), depending on the device's previous state.
\commentAL{TODO Rephrase:} We also note $o^{(d)}_t=1$ (resp. 0) when the user toggled (resp. did not toggle) device $d$ at time $t$.

@@

$$ \forall d \in \mathcal{D},
\begin{cases}
c^{(d)}_t=0 &\text{when } t=0; \\
c^{(d)}_t=c^{(d)}_{t-1} \oplus o^{(d)}_t &\text{when } t>0.
\end{cases}$$

@@

\commentAL{Need to explain how devices go from sequence to the HMM at some point. 
We first define the prediction of behavior from the HMM directly.}

\subsubsection{Predicting future behavior} % (fold)
\label{ssub:predicting_future_behavior}
\commentAL{Why we need to make these predictions is not explained yet.}
Devices want to predict whether they are likely to be connected in the near future.
This information will be publicly available, and allow routes prediction for the upcoming files exchange.

Given the model $\lambda_N$, the current user's location $S_t$ and the current connection state of the devices $C_t$, we can predict the devices connections in $k$ round $C_{t+k}$.

If the device $d$ is connected at time $t$ (resp. disconnected), the probability that $d$ is connected at time $t+k$ is equal to the probability that the user toggles $d$ an even (resp. odd) number of times between round $t+1$ and $t+k$:

$$
P\left[c^{(d)}_{t+k}=1 \mid c^{(d)}_{t}=1 ,S_t \right]= P\left[ \sum\limits_{i\in[\![ 1,k]\!]}o^{(d)}_{t+i} \text{ is even} \mid S_t\right] 
$$

To compute this probability, one thus need to predict the probability that the user toggles $d$ at each round until $t+k$. This is done recursively:

\begin{multline*}
P\left[o^{(d)}_{t+i}=1 \mid S_{t} \right] = \\
\sum\limits_{s \in \mathcal{S}} P\left[ S_{t+1}=s \mid S_t\right] * P\left[o^{(d)}_{t+i}=1 \mid S_{t+1}=s \right]
\end{multline*}

Knowing that the probability of toggling $d$ during the next round is computed as follows:

\begin{multline*}
P\left[o^{(d)}_{t+1}= 1 \mid S_{t} \right]=\\
\sum\limits_{s \in \mathcal{S}} P\left[ S_{t+1}=s \mid S_{t}\right] * P\left[ o^{(d)}_{t+1}=1 \mid S_{t+1}=s\right] 
\end{multline*}

Finally, we enumerate the set $\mathcal{B}^{(k)}_\text{even}$ (resp. $\mathcal{B}^{(k)}_\text{even}$) of sequences of $k$ booleans having an even (resp. odd) number of true values, and compute the probability that device $d$ will be connected in $k$ turns, knowing that it is connected (resp. disconnected) at time $t$:

\begin{multline*}
P\left[c^{(d)}_{t+k}=1 \mid c^{(d)}_{t}=1 ,S_t \right]=\\
\sum\limits_{B \in \mathcal{B}_\text{even}}\sum\limits_{i\in[\![ 1,k]\!]} P\left[o^{(d)}_{t+i}= B\left[i\right] \mid S_{t} \right]
\end{multline*}

\begin{multline*}
P\left[c^{(d)}_{t+k}=1 \mid c^{(d)}_{t}=0 ,S_t \right]=\\
\sum\limits_{B \in \mathcal{B}_\text{odd}}\sum\limits_{i\in[\![ 1,k]\!]} P\left[o^{(d)}_{t+i}= B\left[i\right] \mid S_{t} \right]
\end{multline*}

% \begin{multline}
% P\left[o^{(d)}_{t+1}= 1 \mid S_{t} \right]= P\left[o^{(d)}_{t+i+1}= 1 \mid S_{t+i} \right] = \\
% \sum\limits_{s \in \mathcal{S}} P\left[ S_{t+1}=s \mid S_{t}\right] * P\left[ o^{(d)}_{t+1}=1 \mid S_{t+1}=s\right] \\

% P\left[o^{(d)}_{t+i}=1 \mid S_{t} \right] = \sum\limits_{s \in \mathcal{S}} P\left[ S_{t+1}=s \mid S_t\right] * P\left[o^{(d)}_{t+i}=1 \mid S_{t+1}=s \right]
% \end{multline}

% Using the matrices notations:

% $$
% \begin{cases}
% P\left[o^{(d)}_{t+i+1}=1 \mid S_{t+i} = i\right] = \sum\limits_{j \in \mathcal{S}} a_{ij} * b_{jk}\\

% P\left[o^{(d)}_{t+i}=1 \mid S_{t} = i \right] = \sum\limits_{j \in \mathcal{S}} a_{ij} * P\left[o^{(d)}_{t+i}=1 \mid S_{t+1}=s \right]
% \end{cases}
% $$

% subsubsection predicting_future_behavior (end)
